Learning STRIPS action models with classical planning

### 2017-10-10 doc/icaps-18/strips-learning.tex
starting to add first EVA's comments to the ICAPS18 paper

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{multicol}
\usepackage{arydshln}

@@

\author{Diego Garc\'ia\and Sergio Jim\'enez\and Eva Onaindia\\
{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
{\small Universitat Polit\'ecnica de Val\'encia.}\\
{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
{\small \{,serjice,onaindia\}@dsic.upv.es}}

\maketitle
\begin{abstract}
This paper presents a novel approach for learning STRIPS action models from examples that compiles the inductive learning task into a classical planning task. Our compilation for this learning task is flexible to different amounts of available input knowledge; it accepts partially specified action models and what is more, the input learning examples can range from a set of plans (with their corresponding initial and final states) to only a set of initial and final states where no action is observed.   
\end{abstract}

@@

On the other hand, Machine Learning (ML) techniques are able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}. However, the application of inductive ML techniques for learning planning action models is not straightforward. First, the inputs to ML algorithms usually are sets of finite numeric vectors encoding the value of sets of objects features. The input for learning planning action models traditionally are sets of observations of plan executions (each with possibly different length). Second, the traditional output of off-the-shelf ML techniques is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks). In the case of learning STRIPS action models, the output is not a scalar but a model of the preconditions and the effects of each action that defines the possible state transitions in the given planning domain.

Learning STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of all the states traversed by the example plans. Motivated by recent advances on learning generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating} this paper introduces an innovative approach for learning classical planning action models that (1) it can be defined as a classical planning compilation and (2), it is flexible to different amounts of available input knowledge.

@@

@@

@@

\subsection*{Acknowledgment}
\begin{small}
Diego Garc\'ia is partially supported by the {\it FPU} program funded by the Spanish government. Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
\end{small}

>>>>

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{multicol}
\usepackage{arydshln}
\usetikzlibrary{calc,backgrounds,positioning,fit}

@@

\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
{\small Universitat Polit\'ecnica de Val\'encia.}\\
{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
{\small \{,serjice,onaindia\}@dsic.upv.es}}

\maketitle
\begin{abstract}
This paper presents a novel approach for learning STRIPS action models from examples that compiles this inductive learning task into a classical planning task. Interestingly the proposed compilation is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no actions or intermediate states are observed. What is more, the compilation can reuse previous knowledge (it is able to start from partially specified action models) and can be used to validate whether an observed plan follows a given action model. 
\end{abstract}

@@

On the other hand, Machine Learning (ML) techniques are able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}. However, the application of inductive ML techniques for learning planning action models is not straightforward. First, the inputs to ML algorithms usually are finite vectors encoding the value of a given set of features in a set of objects. The input for learning planning action models traditionally are sets of observations of plan executions (each with a possibly different length). Second, the traditional output of off-the-shelf ML techniques is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks). When learning STRIPS action models, the output is not a scalar but a model of the preconditions and the effects of each action that defines the possible state transitions of a given planning domain.

Learning STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of all the intermediate states traversed by the example plans. Motivated by recent advances on learning generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating} this paper introduces an innovative approach for learning classical planning action models that:
\begin{itemize}
\item Can be defined as a classical planning compilation.
\item It is flexible to different amounts of available input knowledge. The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no actions or intermediate states are observed.
\item Can exploit previous knowledge about the action model in the form of partially specified action models.
\end{itemize}

The paper is organized as follows. The first section presents the classical planning model and its extension to define conditional effects. The second section formalizes the learning task addressed in the paper. The third section describes our approach for addressing the learning of STRIPS action models by compiling this task into a classical planning task. The fourth section evaluates the performance of our approach. Finally the last section discusses  the strengths and weaknesses of our approach and proposing some future work.

@@

\begin{figure}[hbt]
\begin{scriptsize}
\begin{verbatim}
;;; Predicates as defined in PDDL
(on ?o1 - object ?o2 - object) (ontable ?o - object) 
(clear ?o - object) (holding ?o  - object)
(handempty)
\end{verbatim}

\vspace{0.1cm}

\begin{subfigure}{.3\textwidth}
\begin{verbatim}
;;; Plan
0: (unstack A B)
1: (put-down A)
2: (unstack B C)
3: (stack B A)
4: (unstack C D)
5: (stack C B)
6: (pick-up D)
\end{verbatim}
\end{subfigure}%
\begin{subfigure}{.6\textwidth}
\begin{verbatim}
;;; Initial and final states
\end{verbatim}
\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
\node (initA) []{A};
\node (initB) [above=of initA.north]{B};
\node (initC) [above=of initB.north]{C};
\node (initD) [above=of initC.north]{D};

\node (goalD) [right=of initA.south east]{D};
\node (goalC) [right=of initB.south east]{C};
\node (goalB) [right=of initC.south east]{B};
\node (goalA) [right=of initD.south east]{A};
\end{tikzpicture}

\end{subfigure}%
\end{scriptsize}
 \caption{\small Example of a learning task $\Lambda$ with a single learning example that corresponds to invert a tower of four blocks.}
\label{fig:lexample}
\end{figure}

To illustrate this, Figure~\ref{fig:lexample}, shows $\Psi$ and $\Pi=\{\pi_1\}$ for a learning task $\Lambda$ with a single learning example that corresponds to invert a tower of four blocks.

@@

\begin{figure}[hbt]
\begin{footnotesize}
\begin{scriptsize}
\begin{verbatim}
0 : (program_pre_clear_stack_var1)
1 : (program_pre_handempty_stack_)
2 : (program_pre_holding_stack_var2)
3 : (program_pre_on_stack_var1_var1)
4 : (program_pre_on_stack_var1_var2)
5 : (program_pre_on_stack_var2_var1)
6 : (program_pre_on_stack_var2_var2)
7 : (program_pre_ontable_stack_var1)
8 : (program_pre_ontable_stack_var2)
9 : (program_eff_clear_stack_var1)
10 : (program_eff_clear_stack_var2)
11 : (program_eff_handempty_stack_)
12 : (program_eff_holding_stack_var1)
13 : (program_eff_on_stack_var1_var2)
14 : (validate_0)
15 : (unstack a b i1 i2)
16 : (put-down a i2 i3)
17 : (unstack b c i3 i4)
18 : (stack b a i4 i5)
19 : (unstack c d i5 i6)
20 : (stack c b i6 i7)
21 : (pick-up d i7 i8)
22 : (validate_1)
\end{verbatim}
\end{scriptsize}
\end{footnotesize}
 \caption{\small Example of a plan.}
\label{fig:plan}
\end{figure}

@@

\subsection*{Acknowledgment}
\begin{small}
Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government. Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
\end{small}
